---
title: "Comparison of Percent Democratic vote vs the "
output: html_notebook
---

Looking to compare local area unemployment statistics vs who a county voted for in 2016. To see if Democrates vs republicans are experiencing the pandamic differently. It appears to me that we will see a big increase in Democratic unemployment but not as much for republicans. 



```{r}
#Libraries
library(vroom)
library(dplyr)
library(ggplot2)
```


```{r}
#Data from Unemployment 
unemploymentURL = "https://www.bls.gov/web/metro/laucntycur14.txt"
unemployment = vroom(unemploymentURL, delim = "|", skip = 6)
head(unemployment)
names(unemployment)<- c("LAUS_Area","State_Fips","County_Fips","Area_Name","Period","Labor_Force","Employed","Unemployed_People","Unemployed_Rate")
head(unemployment)
unemployment$FIPS = paste0(unemployment$State_Fips,unemployment$County_Fips)


#unemployment$DateEstimate = as.Date(unemployment$Period,"%B-%y")

```
Now we need to import the county data for who a county voted for as president in 2016. data comes from "https://github.com/Deleetdk/USA.county.data/blob/master/data/USA_county_data.RData"


```{r}
load("C:/Users/steve/Downloads/USA_county_data.RData")
USA_county_data #Name of the data from github. 

votes = USA_county_data
head(votes)
votes$FIPS = ifelse(nchar(votes$fips)<5, paste0("0",as.character(votes$fips)), as.character(votes$fips))

votes[votes$FIPS=="21005",]
votes %>%  filter(statecode_prev=="21")



```

Join the 2 datasets together as a left loin with unemployment as the left. Note Alaska has no county data, so is excluded.

```{r}
unemployed_Voters = left_join(unemployment,votes, by="FIPS")
head(unemployed_Voters)

sum(ifelse(unemployed_Voters$Unemployed_Rate>0.1,1,0))

sum(is.na(unemployed_Voters$Unemployed_Rate))
sum(is.na(unemployed_Voters$dem16_frac))
```
Plot the comparison of voted Clinton to Unemployment
```{r}

ggplot(data= unemployed_Voters, aes(x = dem16_frac, y = Unemployed_Rate))+
  geom_point()+
  geom_smooth()+
  labs(title = 'Percent of 2016 Votes for Democrates vs May 2020 Unemployment')+
  xlab( "Percent Democratic vote")+
  ylab(" Percent Unemployment May 2020")+
  theme_bw()
  
  

```

